Networks of financial actors

### 2021-09-23 networks.Rmd
collecting geographical coordinates

---
title: "Networks of financial actors"
author: "Juan Rocha"
date: "6/10/2021"
output: 
  html_notebook:
      theme:
          code_font:
            google: Fira Code
---

@@

```{r}
library(tidyverse)
library(plotly)
library(patchwork)
library(network)

@@

Currently the data set has `r df1 %>% nrow()` companies, in `r df1$country %>% unique() %>% length()` countries, dealing with `r df1$commodity %>% unique() %>% length()` commodities (including `NA`).

The data is very incomplete with respect to country locations

@@

```{r}
library(sna)
net_countries <-  df1 %>% 
    select(company, origin, country) %>% 
    group_by(origin, country) %>% 
    # unique() %>% 
    count(name = "channel") %>% # a channel is acountry * commodities
    filter(!is.na(origin), !is.na(country)) %>% 
    network(directed = TRUE, bipartite = FALSE, matrix.type = "edgelist", ignore.eval=FALSE, loops = FALSE)

@@

```{r}
df1 %>% 
    select(company, origin, country) %>% 
    group_by(origin, country) %>% 
    # unique() %>% 
    count(name = "channel") %>% # a channel is acountry * commodities
    filter(!is.na(origin), !is.na(country)) %>% 
    filter(origin != country) %>% 
    arrange(desc(channel))
```

@@

```{r}
df1 %>% 
    select(company, commodity) %>% 
    group_by(company, commodity) %>% 
    count(name = "geographies") %>% 
    arrange(desc(geographies)) %>% 
    ggplot(aes(company, commodity)) +
    geom_tile(aes(fill = geographies)) +
    scale_fill_viridis_c() +
    theme_light(base_size = 6) +
    theme(axis.text.x = element_text(angle = 90)) 
```

>>>>

---
title: "Networks of financial actors"
author: "Juan Rocha"
date: "Last update `r Sys.Date()`"
output: 
  html_notebook:
      theme:
          code_font:
            google: Fira Code
---

@@

```{r}
library(tidyverse)
library(plotly)
library(patchwork)
library(network)
library(sna)

@@

Currently the data set has `r df1 %>% nrow()` companies, in `r df1$country %>% unique() %>% length()` countries, dealing with `r df1$commodity %>% unique() %>% length()` commodities (including `NA`). Below some overview of the missing values:

@@

```{r}

net_countries <-  df1 %>% 
    select(company, origin, country) %>% 
    group_by(origin, country) %>% 
    # unique() %>% 
    count(name = "channel") %>% # a channel is acountry * commodities
    filter(!is.na(origin), !is.na(country)) %>% 
    network(directed = TRUE, bipartite = FALSE, matrix.type = "edgelist", ignore.eval=FALSE, loops = TRUE)

@@

```{r}
df1 %>% 
    select(company, origin, country) %>% 
    group_by(origin, country) %>% 
    # unique() %>% 
    count(name = "channel") %>% # a channel is a country * commodities
    filter(!is.na(origin), !is.na(country)) %>% 
    filter(origin != country) %>% 
    arrange(desc(channel))
```

@@

```{r fig.width = 6, fig.height = 5}
df1 %>% 
  select(company, commodity) %>% 
  group_by(company, commodity) %>% 
  count(name = "geographies") %>% 
  arrange(desc(geographies)) %>% 
  filter(geographies > 3) %>%  ungroup() %>% 
  mutate(company = as_factor(company)) %>%
  ## the data needs to be ungrouped for fct_reorder to work!
  mutate(company = fct_reorder(
    .f = company, .x = geographies, .fun = sum, .desc = FALSE)) %>% 
  #pull(company) %>% levels()
  ggplot(aes(y = company)) +
  geom_bar(aes(weight = geographies, fill = commodity),
           position = "stack", show.legend = TRUE, alpha = 0.5) +
  scale_fill_discrete(
    "Commodities", guide = guide_legend(
      ncol = 4, title.position = "top", title.hjust = 0.5,
      keywidth = 0.5, keyheight = 0.5
    )
  ) +
  labs(title = "Companies ranking by number of countries and commodities",
       subtitle = "A geography is a combination of `country * commodity` as a proxy of geographical influence of a company.\nOnly companies with >3 geographies are shown", caption = "Data curated by the project, source Orbis database", x = "Geographies", y = "Top companies") +
  theme_light(base_size = 10) + theme(legend.position = c(0.7, 0.1))

Note there is a relatively low diversity of specialization within companies, meaning that a lot of them are putting their eggs on the same basket.

Another key network representation is by companies given the countries where they together operate, and by the common geographies (meaning exporting the same commodity from the same country).

```{r}
df1 %>% 
  select(company, origin, country) %>% 
  group_by(origin, company) %>% 
  # Assumption: if country is NA, it means the company operates in its country of origin:
  mutate(country = case_when(is.na(country) ~ origin, !is.na(country) ~ country)) %>% 
  # unique() %>% 
  count(name = "channel") %>% # a channel is acountry * commodities
  ungroup() %>%
  pivot_wider(values_from = geographies, names_from = companies, values_fill = 0)

   # filter(!is.na(origin), !is.na(company)) %>% 
  network(directed = TRUE, bipartite = FALSE, matrix.type = "edgelist", ignore.eval=FALSE, loops = TRUE)

net_companies %v% "indegree" <- sna::degree(
    net_companies, gmode = "digraph", diag = TRUE, rescale = FALSE, cmode = "indegree")
net_companies %v% "outdegree" <- sna::degree(
    net_companies, gmode = "digraph", diag = TRUE, rescale = FALSE, cmode = "outdegree")
net_companies %v% "degree" <- sna::degree(net_companies, gmode = "digraph", diag = TRUE, rescale = FALSE)

gplot(
    net_companies,
    vertex.col = alpha("orange", 0.25),
    label = network.vertex.names(net_companies),
    label.cex = 0.5, label.col = "grey50",
    vertex.border = 0, 
    vertex.cex = sqrt(0.2 + net_companies %v% "outdegree"),
    label.pos = 5,
    edge.col = alpha("grey50", 0.1),
    edge.lwd = log1p(net_companies %e% "channel"),
    displayisolates = TRUE, pad = 0.5,
    mode = "mds"
    )
